### 2015-06-08 publishing/response-to-reviewers.md
added first manuscript and started revision

PONE-D-15-17915
3D-Printed 24-Well Plate Oxygen Control Device
PLOS ONE

Dear Arum Han and reviewers,

Thank you for reviewing our manuscript submission to PLOS ONE. We have made several improvements and clarifications resulting what we hope is a stronger manuscript ready for publication.

If you would like to make changes to your financial disclosure, please include your updated statement in your cover letter.

In addition to this rebuttal letter a clean revised manuscript is submitted as well as a marked-up copy with tracked changes.

Yours sincerely,

David Eddington, Ph.D.
University of Illinois at Chicago

Authors' Rebuttal to Reviewers

Reviewer #1:
It is a different version of design to control oxygen in cell culture, among some recent efforts. One concern is that the group has published quite a number of papers on oxygen control devices, although it is now through 3d printing. It is important to highlight in front the key improvements.

Authors' rebuttal:

Also, it is important to compare with some other methods related to 3d printing such as that done recently by Kamei K. et al. (Biomed Microdevices. 2015 Apr;17(2):36).

Authors' rebuttal:

"This is the first 3D-printed device that can be functionalized to control oxygen in cell culture" is a statement which is too strong.

Authors' rebuttal:

------------------------------

Reviewer #2:
Currently, little consideration is given to mimicking physiological oxygen concentrations in vitro during cell experiments. Oftentimes cells are cultured at atmospheric oxygen conditions which are higher than what is found in many tissues, including tumor environments. Some systems have been developed to control the oxygen environment in vitro, but these can be cumbersome or costly to implement. 3D printing has emerged as a rapid prototyping method to fabricate devices with high control over structure and therefore is an attractive method to design a microfluidic device capable of controlling oxygen concentrations in culture. Brennan et al. report on a 3D printed control device capable of adjusting oxygen concentrations in a 24-well plate. Although the concept is interesting, the data is not sufficient and additional experiments are needed to improve the manuscript. 
•	Page 3 - In addition to presenting gap distance below insert and well bottom, it would be helpful if the authors included volume of media this space would hold.

Authors' rebuttal:

•	Page 4 – Why was 6 hrs chosen for cell culture? It would seem that the authors could demonstrate a more robust change in expression over the concentrations tested if the experiment was extended to a longer time point, especially considering only the 0% group showed a significant difference in VEGFA expression.

Authors' rebuttal: VEGFA is under the transcriptional activity of HIF-1alpha. HIF protein peaks at 4 hr, and so we looked at a time point slightly delayed. By longer time points, the HIF response is actually suppressed, and many transcripts are not increased with longer time points. See Rexius-Hall et al.s' work.

•	The authors specify concentrations of 21%, 10%, 5%, and 0% were evaluated, however in their Figure 2b, the concentrations appear closer to 13% (not 10%) and 7% (not 5%). Variation of 3% can be considerable in physiological environments and it is unclear if the authors recognize this variation. It is also amplified by the lack of error bars for each group.

Authors' rebuttal:

•	Is sterilization an issue for this device, considering it will be in close contact with biological components?

Authors' rebuttal: